---
title: "R Notebook"
author: paulbuckley
date: 08/03/2022
output: html_document
---


```{r}
library(data.table)
library(dplyr)
library(tidyr)
library(ggpubr)
library(Biostrings)
library(tidyverse)
library(foreach)
library(stringdist)
library(doParallel)

```

## Useful functions

```{r}

# Function to provide a closest match. Used to match HLA Alleles across mixed output styles.
ClosestMatch2 = function(string, stringVector){

  stringVector[amatch(string, stringVector, maxDist=Inf)]

}

```

## Aims
Question: If we look at ratios of binding (agretopicity), stability, and/or fraction of hydrophobicity, do we see any evidence of immune escape?

## Method
- I have alresady predicted binding of these class I CD8T epitoeps and their mutants to 50 class I alleles.
- I read in all information.


# Read in Wuhan binding dataset
- Sample of what the data look like.
```{r}

WUHAN_PEPTIDES_BINDING_SCORES=readRDS("ORIGINAL_PEPTIDES_BINDING_SCORES_V2.rds")
WUHAN_PEPTIDES_BINDING_SCORES %>% head
WUHAN_PEPTIDES_BINDING_SCORES %>% nrow


```

## For 9/10mers of the wuhan originals, what are dominant HLAs?

```{r,dpi=300,fig.height=7}

WUHAN_PEPTIDES_BINDING_SCORES%>% mutate(Length=nchar(Peptide)) %>% filter(Length%in% c(9,10))%>% select(Peptide, Predicted_Binding, Binder,Length) %>% separate_rows_(cols=c("Predicted_Binding", "Binder"), sep=",") %>% distinct() %>% group_by(Predicted_Binding,Binder,Length) %>%mutate(Length = as.character(Length))%>%
    dplyr::summarise(n=n())%>% filter(Binder == "BINDER")%>% arrange(desc(n))%>%
    ggbarplot(x="Predicted_Binding",y="n",fill="Length",position = position_dodge())+coord_flip()

WUHAN_PEPTIDES_BINDING_SCORES%>% mutate(Length=nchar(Peptide)) %>% filter(Length%in% c(9,10))%>% select(Peptide, Predicted_Binding, Binder,Length) %>% separate_rows_(cols=c("Predicted_Binding", "Binder"), sep=",") %>% distinct() %>% filter(Predicted_Binding == "HLA-A0201")%>% group_by(Predicted_Binding,Binder,Length) %>%mutate(Length = as.character(Length))%>%
    dplyr::summarise(n=n())%>% filter(Binder == "BINDER")%>% arrange(desc(n))



```


# Read in Omicron mutanta binding dataset
- Sample of what the data look like.
```{r}

OMICRON_PEPTIDES_BINDING_SCORES=readRDS("OMICRON_PEPTIDES_BINDING_SCORES_V2.rds")
OMICRON_PEPTIDES_BINDING_SCORES %>% head
OMICRON_PEPTIDES_BINDING_SCORES %>% nrow


```


## For 9/10mers of the omicron mutants, what are dominant HLAs?
- HLA-Bs and HLA-Cs.
- only 7 bind A0201

```{r,dpi=300,fig.height=7}

OMICRON_PEPTIDES_BINDING_SCORES%>% mutate(Length=nchar(VariantAlignment)) %>% filter(Length%in% c(9,10))%>% select(VariantAlignment, MT_Predicted_Binding, MT_Binder,Length) %>% separate_rows_(cols=c("MT_Predicted_Binding", "MT_Binder"), sep=",") %>% distinct() %>% group_by(MT_Predicted_Binding,MT_Binder,Length) %>%mutate(Length = as.character(Length))%>%
    dplyr::summarise(n=n())%>% filter(MT_Binder == "BINDER")%>% arrange(desc(n))%>%
    ggbarplot(x="MT_Predicted_Binding",y="n",fill="Length",position = position_dodge())+coord_flip()

OMICRON_PEPTIDES_BINDING_SCORES%>% mutate(Length=nchar(VariantAlignment)) %>% filter(Length%in% c(9,10))%>% select(VariantAlignment, MT_Predicted_Binding, MT_Binder,Length) %>% separate_rows_(cols=c("MT_Predicted_Binding", "MT_Binder"), sep=",") %>% distinct() %>% filter(MT_Predicted_Binding == "HLA-A0201")%>% group_by(MT_Predicted_Binding,MT_Binder,Length) %>%mutate(Length = as.character(Length))%>%
    dplyr::summarise(n=n())%>% filter(MT_Binder == "BINDER")%>% arrange(desc(n))



```


# Read in mutations

```{r}
MUTATIONS = readRDS("OMICRON_EPITOPE_MUTATIONS.rds")
MUTATIONS %>% mutate(Length = nchar(Peptide))%>% filter(Protein == 'surface glycoprotein') %>% filter(Length== 9)%>% nrow
#MUTATIONS=MUTATIONS %>% group_by(Peptide, VariantAlignment,start_pos,Mutation) %>% dplyr::summarise(Protein = paste0(Protein, collapse = ","))

```

# join all

```{r}

MUTATIONS %>% inner_join(WUHAN_PEPTIDES_BINDING_SCORES)%>% nrow
MUTATIONS %>% inner_join(OMICRON_PEPTIDES_BINDING_SCORES) %>% nrow

FULL_MUTATIONS_DT=MUTATIONS %>% inner_join(WUHAN_PEPTIDES_BINDING_SCORES) %>% inner_join(OMICRON_PEPTIDES_BINDING_SCORES)
#FULL_MUTATIONS_DT %>% head %>% DT::datatable()

#saveRDS(FULL_MUTATIONS_DT,file="FULL_MUTATIONS_BINDING.rds")
# mUTATED+HOMOLOGOUS oeptide
FULL_MUTATIONS_DT %>% filter(Peptide %in% "WLLWPVTLA")


FULL_MUTATIONS_DT %>% select(Peptide, VariantAlignment) %>% distinct()
FULL_MUTATIONS_DT %>% nrow

DATA_FOR_ANALYSIS=FULL_MUTATIONS_DT %>% select(Peptide, VariantAlignment, Predicted_Binding, EL_Rank, MT_EL_Rank, Binder,MT_Binder) %>% separate_rows_(cols = c("Predicted_Binding","EL_Rank","MT_EL_Rank","Binder","MT_Binder"),sep=",")


DATA_FOR_ANALYSIS %>% nrow

```

##  Loss / gain plots for HLA A/B peptides

- A/B asupertypes were mapped as of https://bmcimmunol.biomedcentral.com/articles/10.1186/1471-2172-9-1#Abs1
- HLA C were inferred by using the representitive
- Try HLA A/B first
- Problem is 1) low sample number, and 2) not sure if just essentially plotting the distribution of binding in thw dataset. Some HLA alleles (captued by supertypes), will bind these peptides more, thus potentially captured in loss/gain more.
- Need to think through more.

```{r}

AB_SUPERTYPES=fread("HLA_AB_SUPERTYPES.csv") %>% mutate(Allele = gsub("\\*","",Allele))%>% mutate(Allele = paste0("HLA-",Allele))%>% dplyr::rename(HLA_Allele = Allele)

DATA_FOR_ANALYSIS%>% dplyr::rename(HLA_Allele=Predicted_Binding)%>% filter(HLA_Allele %in% grep("HLA-A|HLA-B",HLA_Allele,value = T))%>% nrow

HLA_A_B_DATA_FOR_ANALYSIS =DATA_FOR_ANALYSIS%>% dplyr::rename(HLA_Allele=Predicted_Binding)%>% filter(HLA_Allele %in% grep("HLA-A|HLA-B",HLA_Allele,value = T))%>% mutate(EL_Rank = as.numeric(EL_Rank), MT_EL_Rank=as.numeric(MT_EL_Rank))

HLA_A_B_DATA_FOR_ANALYSIS=HLA_A_B_DATA_FOR_ANALYSIS%>% mutate(LossGain = case_when(
        (Binder == "BINDER" & MT_Binder == "NONBINDER") ~ "Loss",
        (Binder == "NONBINDER" & MT_Binder == "NONBINDER") ~ "Same_NonBinder",
        (Binder == "NONBINDER" & MT_Binder == "BINDER") ~ "Gain",
        (Binder == "BINDER" & MT_Binder == "BINDER") ~ "Same_Binder"
))

HLA_A_B_DATA_FOR_ANALYSIS %>% filter(LossGain %in% c("Loss","Gain"))%>%group_by(HLA_Allele, LossGain) %>% dplyr::summarise(n=n())%>% inner_join(AB_SUPERTYPES)%>%
        ggboxplot(x="LossGain",y="n",facet.by = "Supertype")+stat_compare_means(label="p.signif")


```

## Agretopicities grouped by sipertype
- HLA C inferred

```{r}

DATA_FOR_ANALYSIS=FULL_MUTATIONS_DT %>% select(Peptide, VariantAlignment, Predicted_Binding, nM_41, MT_nM_41, Binder,MT_Binder) %>% separate_rows_(cols = c("Predicted_Binding","nM_41","MT_nM_41","Binder","MT_Binder"),sep=",")%>% mutate(nM_41 = as.numeric(nM_41), MT_nM_41=as.numeric(MT_nM_41))%>% mutate(Agretopicity = MT_nM_41/nM_41)

#DATA_FOR_ANALYSIS %>% dplyr::rename(HLA_Allele=Predicted_Binding)%>% filter(HLA_Allele %in% grep("HLA-A|HLA-B",HLA_Allele,value = T)) %>% inner_join(AB_SUPERTYPES)%>%
#        ggboxplot(x="Supertype",y="Agretopicity")+scale_y_log10()+rotate_x_text()
# without hlac
HLA_A_B_AGRETOPICITY=DATA_FOR_ANALYSIS %>% dplyr::rename(HLA_Allele=Predicted_Binding)%>% filter(HLA_Allele %in% grep("HLA-A|HLA-B",HLA_Allele,value = T)) %>% inner_join(AB_SUPERTYPES) %>%
filter(! (Binder == "NONBINDER" & MT_Binder == 'NONBINDER'))

HLA_CDATASET_AGRETOPICITY=DATA_FOR_ANALYSIS %>% dplyr::rename(HLA_Allele=Predicted_Binding)%>% filter(!HLA_Allele %in% grep("HLA-A|HLA-B",HLA_Allele,value = T))%>% mutate(Supertype = str_extract(HLA_Allele,"HLA-C[0-9][0-9]"))%>%
        mutate(Supertype = gsub("HLA\\-","",Supertype))%>%
filter(! (Binder == "NONBINDER" & MT_Binder == 'NONBINDER'))

HLA_A_B_AGRETOPICITY %>% rbind(HLA_CDATASET_AGRETOPICITY)%>%
        ggboxplot(x="Supertype",y="Agretopicity")+scale_y_log10()+geom_hline(yintercept = 1,linetype="dashed")+rotate_x_text()


```




```{r,dpi=300, fig.width = 10}
HLA_A_B_AGRETOPICITY %>% rbind(HLA_CDATASET_AGRETOPICITY)%>%arrange((HLA_Allele))%>%
        ggboxplot(x="HLA_Allele",y="Agretopicity",add="jitter")+scale_y_log10()+geom_hline(yintercept = 1,linetype="dashed")+rotate_x_text()

```



```{r,dpi=300, fig.width  = 11, fig.height = 7}


HLA_A_B_AGRETOPICITY %>% rbind(HLA_CDATASET_AGRETOPICITY) %>%
        ggpaired(cond1 = "nM_41", cond2 = "MT_nM_41",line.color = "gray")+facet_wrap(~Supertype,ncol=9)+scale_y_log10()+stat_compare_means(label = "p.signif")+rotate_x_text(angle=45)+ylab("Binding Affinity (nM)")
```



## spike only?

```{r,dpi=300}

DATA_FOR_ANALYSIS=FULL_MUTATIONS_DT %>% filter(Protein == "surface glycoprotein") %>% select(Peptide, VariantAlignment, Predicted_Binding, nM_41, MT_nM_41, Binder,MT_Binder) %>% separate_rows_(cols = c("Predicted_Binding","nM_41","MT_nM_41","Binder","MT_Binder"),sep=",")%>% mutate(nM_41 = as.numeric(nM_41), MT_nM_41=as.numeric(MT_nM_41))%>% mutate(Agretopicity = MT_nM_41/nM_41)

DATA_FOR_ANALYSIS %>% dplyr::rename(HLA_Allele=Predicted_Binding)%>% filter(HLA_Allele %in% grep("HLA-A|HLA-B",HLA_Allele,value = T)) %>% inner_join(AB_SUPERTYPES)%>%
        ggboxplot(x="Supertype",y="Agretopicity")+scale_y_log10()+rotate_x_text()
# without hlac
HLA_A_B_AGRETOPICITY=DATA_FOR_ANALYSIS %>% dplyr::rename(HLA_Allele=Predicted_Binding)%>% filter(HLA_Allele %in% grep("HLA-A|HLA-B",HLA_Allele,value = T)) %>% inner_join(AB_SUPERTYPES) %>%
filter(! (Binder == "NONBINDER" & MT_Binder == 'NONBINDER'))

HLA_CDATASET_AGRETOPICITY=DATA_FOR_ANALYSIS %>% dplyr::rename(HLA_Allele=Predicted_Binding)%>% filter(!HLA_Allele %in% grep("HLA-A|HLA-B",HLA_Allele,value = T))%>% mutate(Supertype = str_extract(HLA_Allele,"HLA-C[0-9][0-9]"))%>%
        mutate(Supertype = gsub("HLA\\-","",Supertype))%>%
filter(! (Binder == "NONBINDER" & MT_Binder == 'NONBINDER'))

HLA_A_B_AGRETOPICITY %>% rbind(HLA_CDATASET_AGRETOPICITY)%>%
        ggboxplot(x="Supertype",y="Agretopicity")+scale_y_log10()+geom_hline(yintercept = 1,linetype="dashed")+rotate_x_text()


```

```{r,dpi=300, fig.width = 10}
HLA_A_B_AGRETOPICITY %>% rbind(HLA_CDATASET_AGRETOPICITY)%>%arrange((HLA_Allele))%>%
        ggboxplot(x="HLA_Allele",y="Agretopicity",add="jitter")+scale_y_log10()+geom_hline(yintercept = 1,linetype="dashed")+rotate_x_text()

```

```{r, dpi = 300, fig.width = 10}
HLA_A_B_AGRETOPICITY %>% rbind(HLA_CDATASET_AGRETOPICITY) %>% dplyr::rename(WT_Binder=Binder)%>%
        pivot_longer(cols = c(WT_Binder, MT_Binder))%>%
        group_by(Supertype, name,value) %>% dplyr::summarise(n=n())%>% filter(value == "BINDER")%>% ungroup%>% arrange(desc(n))%>% mutate(Supertype = forcats::fct_inorder(Supertype))%>%
        ggbarplot(x="Supertype",y="n",fill="name",position = position_dodge())+rotate_x_text(angle=90)+ylab("# peptides w/ mutations in Omicron predicted to bind")

```


# spike only


```{r,dpi=300, fig.width  = 11, fig.height = 7}


HLA_A_B_AGRETOPICITY %>% rbind(HLA_CDATASET_AGRETOPICITY) %>%
        ggpaired(cond1 = "nM_41", cond2 = "MT_nM_41",line.color = "gray")+facet_wrap(~Supertype,ncol=9)+scale_y_log10()+stat_compare_means(label = "p.signif")+rotate_x_text(angle=45)+ylab("Binding Affinity (nM)")

BA1_SUPERTYPE_BOXPLT=HLA_A_B_AGRETOPICITY %>% rbind(HLA_CDATASET_AGRETOPICITY)%>% filter(Supertype %in% c("A02","A03","B07","C01"))%>%
        dplyr::rename(Wuhan=nM_41, Omicron = MT_nM_41)%>%
        ggpaired(cond1 = "Wuhan", cond2 = "Omicron",line.color = "gray")+facet_wrap(~Supertype,nrow=1)+theme_pubr(base_size = 16)+scale_y_log10()+stat_compare_means(label = "p.signif",label.x.npc = "center")+rotate_x_text(angle=45)+ylab("Log10 Binding Affinity (nM)")

BA1_SUPERTYPE_BOXPLT
saveRDS(BA1_SUPERTYPE_BOXPLT,file="BA1_SUPERTYPE_BOXPLT.rds")

```


```{r}

HLA_A_B_AGRETOPICITY %>% rbind(HLA_CDATASET_AGRETOPICITY)%>% filter(Supertype == 'B07') %>% arrange(desc(Agretopicity))

HLA_A_B_AGRETOPICITY %>% rbind(HLA_CDATASET_AGRETOPICITY) %>% dplyr::rename(WT_Binder=Binder)%>%
        pivot_longer(cols = c(WT_Binder, MT_Binder))%>%
        group_by(Supertype, name,value) %>% dplyr::summarise(n=n())%>% filter(value == "BINDER")%>% ungroup%>% arrange(desc(n))%>%filter(Supertype == 'B07')%>% arrange((n))%>%
  mutate(pct_change = (n/lead(n) -1) * 100)

HLA_A_B_AGRETOPICITY %>% rbind(HLA_CDATASET_AGRETOPICITY) %>% dplyr::rename(WT_Binder=Binder)%>%
        pivot_longer(cols = c(WT_Binder, MT_Binder))%>%
        group_by(Supertype, name,value) %>% dplyr::summarise(n=n())%>% filter(value == "BINDER")%>% ungroup%>% arrange(desc(n))%>%filter(Supertype == 'A01')%>% arrange((n))%>%
  mutate(pct_change = (n/lead(n) -1) * 100)




```


```{r, dpi = 300, fig.width = 10}
HLA_A_B_AGRETOPICITY %>% rbind(HLA_CDATASET_AGRETOPICITY) %>% dplyr::rename(WT_Binder=Binder)%>%
        pivot_longer(cols = c(WT_Binder, MT_Binder))%>%
        group_by(Supertype, name,value) %>% dplyr::summarise(n=n())%>% filter(value == "BINDER")%>% ungroup%>% arrange(desc(n))%>% mutate(Supertype = forcats::fct_inorder(Supertype))%>%
        ggbarplot(x="Supertype",y="n",fill="name",position = position_dodge())+rotate_x_text(angle=90)+ylab("# peptides w/ mutations in Omicron predicted to bind")

```




